---
title: "University student analysis"
author: "Anna Quaglieri"
date: "28/05/2019"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    fig_caption: yes
    number_sections: yes
    toc_float: yes
    theme: cosmo
  github_document:
    toc: yes
    toc_depth: 4
---

# Setup

```{r setup,echo = FALSE,message=FALSE, warning=FALSE}
library(tidyverse)
library(Hmisc)
library(MASS)
library(broom)
library(cowplot)
library(MKmisc)
library(plotROC)
library(ggrepel)

knitr::opts_chunk$set(eval = TRUE,echo = TRUE,tidy = TRUE,message = FALSE,include = TRUE,warning = FALSE)
```



```{r read-data}
assess <- read_csv("anonymisedData/assessments.csv")
course <- read_csv("anonymisedData/courses.csv")
student.assess <- read_csv("anonymisedData/studentAssessment.csv")
student.info <- read_csv("anonymisedData/studentInfo.csv")
student.regio <- read_csv("anonymisedData/studentRegistration.csv")
student.vle <- read_csv("anonymisedData/studentVle.csv")
vle <- read_csv("anonymisedData/vle.csv")
```


```{r describe-data}
table(assess$code_module) # 7 modules

glimpse(assess)
glimpse(course)
glimpse(student.assess)
glimpse(student.info) # students demoraphics + final results for all students
glimpse(student.regio) # student registration details
glimpse(student.vle) # interaction with virtual learning envir
glimpse(vle)
```

# Data in the cohort

## Assessments

There are 7 modules (corses) and for each module there can be 3 assessment types (not all present in every module)

```{r assess-describe}
table(assess$assessment_type,assess$code_module)
```

## Students

```{r modules-per-student}
n.modules <- student.info %>%
  group_by(id_student) %>%
  summarise(len.modules = length(code_module))
```

* There is a total of `r length(unique(student.info$id_student))` students enrolled and a student attended a minimum of `r min(n.modules$len.modules)` to a maximum of `r max(n.modules$len.modules)`. 

* There are demographics data and assessment results for all students that registered in a module. 

```{r info-check}
sum(!(student.assess$id_student %in% student.info$id_student))
sum(!(student.info$id_student %in% student.assess$id_student))

sum(!(student.info$id_student %in% student.assess$id_student))

sum(!(student.info$id_student %in% student.regio$id_student))
```

* In theory, if a student also has an unregistration date they should not have a final score for an exam for a module. Indeed, Students who unregistered have `Withdrawal` as the value of the `final_result` column in the `studentInfo.csv` file. However, this is not always the case and I will clean up students with a different `final_result` entry.

* Assessment are submitted throghout the module (different dates) and if a student does not submit an assessment then there is no record. 

* There are `r length(unique(student.assess$id_student[!is.na(student.assess$score)]))` students with at least an assessment available with a valid score. 

* All student with an assessment have demographics available.

* `r length(unique(student.info$id_student[!(student.info$id_student %in% student.assess$id_student)]))` students do not have an assessment available.

* There are `r length(unique(student.info$id_student[!(student.info$id_student %in% student.vle$id_student)]))` students who did not interact with the VLE

* There are `r length(unique(student.assess$id_student[!(student.assess$id_student %in% student.vle$id_student)]))` students with an assessment did not interact with the VLE.

```{r}
sum(!(student.info$id_student %in% student.vle$id_student))
sum(!(student.assess$id_student %in% student.vle$id_student))
```

* A student might have several entries in student.info but not all assignment
* This means that if I merge the assignment score with student infos, I will have some studnet.infos entries with NAs for the score. 
* Also some students have NAs for their scores in the assessment data

# Merge datasets

## VLE

```{r merge-vle}
vle.combine <- vle %>%
left_join(student.vle) %>%
  dplyr::rename(date.vle = date)
```

## Assessment infos

```{r merge-assessment}
# Connect assessment types scores with scores
ass <- student.assess %>%
  left_join(assess)

summary(ass$score)
table(sign(ass$date_submitted),ass$code_module,ass$code_presentation)
table(sign(ass$date_submitted),ass$assessment_type)
ass$date.sign <- sign(ass$date_submitted)
ggplot(ass,aes(x = factor(date.sign), y = score, fill=code_module)) + geom_boxplot()

ass <- ass %>%
  filter(!is.na(score))
```

* There are `r sum(is.na(ass$score))` NAs assessment scores and `r sum(sign(ass$date_submitted) == "-1")` assignments were submitted before the start of the course. Given the large number and my limited knolwledge about the dataset I will keep them in the study. They might be entry tests. However, I will remove the NAs assessment score. 

* On the online resource it says:
date – information about the final submission date of the assessment calculated as the number of days SINCE the start of the module-presentation. The starting date of the presentation has number 0 (zero). How can it be negative??

* Still after filtering this leaves me with `r dim(ass)[1]` assessments across `r length(unique(ass$id_student))` students. 

## Students demographics

```{r merge-student-infos}
# Now I know which student what id assessment did.
combine <- student.info %>% # code_module , code_pres, id_student, demo, final score.
  left_join(student.regio) %>% 
    left_join(course) 
```

# Create combined features: assignment score and number of clicks

```{r create-features}
summary(ass$score)
ass.total <- ass %>%
  group_by(code_module,code_presentation,id_student) %>%
  summarise(comb.score = mean(score))

click.total <- vle.combine %>%
  group_by(code_module,code_presentation,id_student) %>%
  summarise(total.clicks = sum(sum_click))

# Merge assessment and clicks
click.score <- ass.total %>%
  left_join(click.total)
```


# Filter students, modify variables

In this first analysis I am interested in understanding what are the factors that influence withdrawal (1 = withdrawl, 0 = PASS/FAIL/Distinction) and Success (1 = PASS/Distinction, 0 = FAIL). The factors influencing withdrawal could be very different from the factors leading to failure, e.g. external factors versus student's capabilities.

I will create a total weigthed score for the assigment and exams scores available in each module. Some students, if they withdrew do not have score for a module so there will be NAs once combining this meta-score with student infos. I also created a total score for the number of clicks in the VLE done in one module in each semester. I will also create a variable with total number of students in each module. 


## Unregistered/Registered students

All students who unregistered during the course could only be `Withdrawn` as final result since they left the course.  
However, it appears that some students that never `unregistered` still withdrawn from the module. 
Also, Some students never registered but still they unregistered or they are in course??

```{r keep-registered-students}
combine$Unregistered <- ifelse(is.na(combine$date_unregistration),"in course","unregistered")
combine$Registered <- ifelse(is.na(combine$date_registration),"not registered","registered") 

sum.reg <- combine %>%
  group_by(Registered,Unregistered,final_result) %>%
    summarise(n.students = length(unique(id_student)))

knitr::kable(sum.reg,caption = "Number of students by registration date, unregistration date and final result.")
```


```{r filter-unreg}
combine <- combine %>%
  filter(!(Registered %in% "not registered"),
         !(Unregistered  %in% "unregistered" & Registered %in% "registered" & final_result %in% "Fail")) 
```

We have now `r length(unique(combine$id_student))` students.

```{r merge-with-ass-clicks}
dim(combine)
combine.feat <- combine %>%
  left_join(click.score)
dim(combine.feat)
```


## Modify some variables: categorical to numeric

```{r nas-in-variables}
nas <- apply(combine.feat,2,function(x) sum(is.na(x)))
nas
```


```{r recode-vars}
table(combine.feat$highest_education)
table(combine.feat$num_of_prev_attempts)

combine.feat <- combine.feat %>%
  mutate(withdraw = ifelse(final_result %in% "Withdrawn",1,0),
         imd_band_num = as.numeric(as.factor(imd_band)),
         highest_education = factor(highest_education, levels=c("No Formal quals","Lower Than A Level", "A Level or Equivalent","HE Qualification","Post Graduate Qualification")),
         num_of_prev_attempts_dich = ifelse(num_of_prev_attempts > 0,">0","0"),
         year = str_remove(code_presentation, pattern = "B|J"),
         semester = str_extract(code_presentation, pattern = "B|J"),
         code_presentation = factor(code_presentation, levels = c("2013B","2014B","2013J","2014J")))
```


# Filter multiple entries per student

* There are students who attempted the same module multiple times and students who attempted different modules. The replicates account for roughly 12% of the entries. To avoid more complicated random intercept models, if the same student attempted the same module twice then I will keep the entry with the highest number of  num_of_prev_attempts because this information should include information about previous attempts. 
* I will discard randomly all students who attempted different modules. If these aren't removed, the entries won't be independent and it would require introducing a mixed effect model with random intercept for each subject. The model itself is a logistic model with many independent variables and fitting a mixed model could be quite complicated and the majority of the students anyway only have one entry. 

```{r study-replicate-students}
repl <- names(table(combine.feat$id_student)[table(combine.feat$id_student) > 1])
repl.student <- combine.feat[combine.feat$id_student %in% repl,]

repl.student.sum <- repl.student %>%
  group_by(id_student) %>%
  summarise(n.modules = length(unique(code_module)),
            n.results = length(unique(final_result)),
            n.repl = length(code_module))

ggplot(repl.student.sum,aes(x=factor(n.modules),fill = factor(n.results))) + geom_bar(position = "dodge") + facet_wrap(~n.repl) +
  theme_bw()
```


```{r remove-duplicate-students}
student.to.keep <- combine.feat %>%
  group_by(code_module,id_student) %>%
  summarise(code_presentation = code_presentation[which.max(num_of_prev_attempts)]) %>%
  unite(key.student.module, code_module,id_student,code_presentation,sep="-",remove=FALSE)

combine.feat.unique <- combine.feat %>%
  unite(key.student.module, code_module,id_student,code_presentation,sep="-",remove=FALSE) %>%
  filter(key.student.module %in% student.to.keep$key.student.module )

combine.feat.unique <- combine.feat.unique[!duplicated(combine.feat.unique$id_student),]

nrow(combine.feat.unique)/nrow(combine.feat) # removed 12% of the initial data
```


```{r add-nstudent-per-module}
# Total number of students attending a course - includign students who attempted different modules
nstudents <- combine.feat %>%
  group_by(code_module,code_presentation) %>%
  summarise(n.students = length(unique(id_student)))

combine.feat.unique <- combine.feat.unique %>%
  left_join(nstudents)
```


```{r table-dataset-unique}
sum.reg <- combine.feat.unique %>%
  group_by(Registered,Unregistered,final_result) %>%
    summarise(n.students = length(unique(id_student)))
colnames(sum.reg) <- c("Registered","Unregistered","Final result","Number of students")

knitr::kable(sum.reg,caption = "Number of students by registration date, unregistration date and final result.")
```


# Combine VLE clicks data

In 2014 there was an increased activity on the VLe which is probably why it helped improving success. 

```{r vle-type-frequency}
vle.combine.sum <- vle.combine %>%
  group_by(code_module,code_presentation,activity_type,id_student) %>%
  summarise(sum_click = sum(sum_click)) %>%
  unite(key.student.module, code_module,id_student,code_presentation,sep="-",remove=FALSE) %>%
  filter(key.student.module %in% student.to.keep$key.student.module ) %>%
 mutate( year = str_remove(code_presentation, pattern = "B|J"),
         semester = str_extract(code_presentation, pattern = "B|J")) %>%
  left_join(nstudents)

nstudents.year.semester <- combine.feat %>%
  group_by(year,semester) %>%
  summarise(n.students = length(unique(id_student)))

vle.combine.sum.sum <- vle.combine.sum %>%
  group_by(activity_type,year,semester) %>%
  summarise(n.clicks = sum(sum_click)) %>%
  ungroup() %>%
  left_join(nstudents.year.semester) %>%
  mutate(n.clicks.std = n.clicks/n.students) %>%
    mutate(activity_type =  fct_reorder(activity_type,n.clicks.std))

ggplot(vle.combine.sum.sum,aes(x = activity_type, y = n.clicks.std, fill = year)) + geom_bar(stat="identity",position="dodge") +facet_wrap(~semester) + coord_flip() + theme_bw()
```

* AAA has got very little students

```{r top5-activities}
vle.combine.top5 <- vle.combine.sum %>%
  filter(activity_type %in% c("oucontent","forumng","homepage","quiz","subpage"))

vle.combine.sum.sum <- vle.combine.top5 %>%
  group_by(activity_type,semester,code_module,n.students) %>%
  summarise(n.clicks = sum(sum_click)/unique(n.students)) %>%
  unite(key , activity_type,semester,code_module,sep="-",remove=FALSE) %>%
  ungroup() %>%
  mutate(activity_type =  fct_reorder(activity_type,n.clicks))


ggplot(vle.combine.sum.sum,aes(x = code_module, y = n.clicks, fill = semester)) + geom_bar(stat="identity",position="dodge",colour="white") +facet_wrap(~activity_type) +  theme_bw() +coord_flip()
```


```{r click-perstudent-topactivity}
# Sum of clicks done by each student per activity type
click.students.activity <- vle.combine.sum %>%
    filter(activity_type %in% c("oucontent","forumng","homepage","quiz","subpage","resource","ouwiki")) %>%
  group_by(code_module,year,semester,activity_type,id_student) %>%
  summarise(n.clicks.act = sum(sum_click))

n.students.vle <- click.students.activity %>%
  group_by(code_module,year,semester) %>%
  summarise(n.students.clicks = length(unique(id_student)))

click.students.activity <- click.students.activity %>%
  left_join(n.students.vle) %>%
  mutate(clicks.per.student = n.clicks.act/n.students.clicks)
```


# Describe modules and students attending each module

* Across two year there were two presentations in each year
* Module AAA is the one with less student, probably aroun 400. Module CCC in 2014 is the one attracting mode students
* There seem to be a semester effect in the number students, with more students enrolling in the J part of the year

```{r affluence-to-courses, fig.cap="Number of students enrolled in each module across years and presentations. The figure included students who enrolled multiple time (same or different module). "}
sum.students <- combine.feat %>%
  group_by(code_presentation,code_module) %>%
  summarise(n.students = length(unique(id_student))) %>%
  ungroup() %>%
  mutate(year = str_remove(code_presentation, pattern = "B|J"),
         semester = str_extract(code_presentation, pattern = "B|J"),
         code_presentation = factor(code_presentation, levels = c("2013B","2014B","2013J","2014J"))) %>%
  dplyr::rename(`Code presentation` = code_presentation)


# There seem to be a semester effect in the number students, with more students enrolling in the J 
ggplot(sum.students, aes(x = code_module, y = n.students, fill = `Code presentation`)) + 
  geom_bar(stat = "identity",position = "dodge",colour="white") + theme_bw() + ggtitle("Number of students enrolled in each module") + theme(legend.position = "bottom") + labs(x  = "Code of module", y = "Number of student enrolled")
```


```{r affluence-to-courses-year}
# There doesn't seem to be a year effect
ggplot(sum.students, aes(x = code_module, y = n.students, fill = year)) + 
  facet_wrap(~ semester) +
  geom_bar(stat = "identity",position = "dodge",colour="white") + theme_bw()+ ggtitle("Number of students enrolled in each module")
```

* The score does not look super discriminant. Am I using it correctly? 

```{r score-by-finalresul}
ggplot(combine.feat.unique, aes(x = comb.score, fill = final_result))  + geom_density() + theme_bw()
```

# Exploratory: factors that influence dropout and success

* People that withdraw tend to have a slightly lower socio economic status

* A lot of region are around the mean withdrawal level. I will take the south west region as reference. 


```{r exploratory-imd_band_num-region}
# Poeple that withdraw tend to have a slightly lower socio economic statis
ggplot(combine.feat.unique,aes(x = factor(withdraw), y = imd_band_num)) + geom_boxplot() + facet_wrap(~semester) + theme_bw()

# Scotland, wales and ireland are the ones with lower levels of widthdrawal
sum.region.with <- combine.feat.unique %>%
  group_by(region) %>%
  summarise(imd.mean = mean(imd_band_num,na.rm = TRUE),
            sem.imd.mean = sd(imd_band_num,na.rm = TRUE)/sqrt(length(imd_band_num[!is.na(imd_band_num)])),
            
            with.mean = mean(withdraw),
            sem.with.mean = sd(withdraw)/sqrt(length(withdraw))) %>%
  mutate(region = fct_reorder(region,with.mean))

ggplot(sum.region.with,aes(x = with.mean, y = imd.mean,colour=region)) + geom_point(stat = "identity")  + theme_bw() +
  geom_errorbar(aes(ymin = imd.mean - sem.imd.mean, 
                    ymax = imd.mean + sem.imd.mean)) +
  geom_errorbarh(aes( xmin = with.mean - sem.with.mean, 
                    xmax = with.mean + sem.with.mean)) +
  geom_label_repel(aes(label = region)) +
  geom_hline(yintercept = mean(combine.feat.unique$imd_band_num,na.rm = TRUE),linetype="dotted") + ylim(c(3.5,7.5)) +
   geom_vline(xintercept = mean(combine.feat.unique$withdraw,na.rm = TRUE),linetype="dotted") 



```

* I will use module BBB as reference for the module, it's around the mean of withdrawal

* The module with higheest prop of withdrawal are CCC/DDD and the ones with less withdrawal are AAA and GGG. The socio econoic area does not seem to be too heterogenous apart from module AAA.

```{r exploratory-imd_band_num-module}
sum.region <- combine.feat.unique %>%
  group_by(code_module,semester,year) %>%
  summarise(imd.mean = mean(imd_band_num,na.rm = TRUE),
            sem.imd.mean = sd(imd_band_num,na.rm = TRUE)/sqrt(length(imd_band_num[!is.na(imd_band_num)])),
            
            with.mean = mean(withdraw),
            sem.with.mean = sd(withdraw)/sqrt(length(withdraw))) %>%
  ungroup() %>%
  mutate(code_module = fct_reorder(code_module,with.mean))

ggplot(sum.region,aes(x = with.mean, y = imd.mean,colour=code_module,shape=year)) + geom_point(stat = "identity",size=2)  + theme_bw() +
  geom_errorbar(aes(ymin = imd.mean - sem.imd.mean, 
                    ymax = imd.mean + sem.imd.mean)) +
  geom_errorbarh(aes( xmin = with.mean - sem.with.mean, 
                    xmax = with.mean + sem.with.mean)) +
  geom_label_repel(aes(label = code_module)) + facet_wrap(~semester) +
  geom_hline(yintercept = mean(combine.feat.unique$imd_band_num,na.rm = TRUE),linetype="dotted") + ylim(c(3.5,7.5))+
   geom_vline(xintercept = mean(combine.feat.unique$withdraw,na.rm = TRUE),linetype="dotted") 
```



```{r exploratory-with-module}
sum.region <- combine.feat.unique %>%
  group_by(code_module,semester,year) %>%
  summarise(with.mean = mean(withdraw),
            sem.with.mean = sd(withdraw)/sqrt(length(withdraw))) %>%
  ungroup() %>%
  mutate(code_module = fct_reorder(code_module,with.mean))

ggplot(sum.region,aes(x = code_module, y = with.mean,colour=code_module,shape=year)) + geom_point(stat = "identity",size=2)  + theme_bw() +
  geom_errorbar(aes(ymin = with.mean - sem.with.mean, 
                    ymax = with.mean + sem.with.mean),width=0.5) + facet_wrap(~semester) +
  geom_hline(yintercept = mean(combine.feat.unique$withdraw,na.rm = TRUE),linetype="dotted")
```

## Exploratory: dropout/success by socio-economical factors

```{r define-pass-fail}
combine.feat.unique <- combine.feat.unique %>%
  mutate(pass_fail = ifelse(final_result %in% c("Fail","Withdrawn"),0,1)) 
```


```{r exploratory-withdraw-pass-fail}
# IMD bands 
imd.band <- unique(combine.feat.unique[,c("imd_band","imd_band_num")]) %>%
  filter(!is.na(imd_band)) %>%
  mutate(imd_band = ifelse(imd_band %in% "10-20","10-20%",imd_band),
         imd.mean = imd_band_num)

# Scotland, wales and ireland are the ones with lower levels of widthdrawal
sum.region.pass.fail <- combine.feat.unique %>%
  group_by(region) %>%
  summarise(imd.mean = round(mean(imd_band_num,na.rm = TRUE)),
            sem.imd.mean = sd(imd_band_num,na.rm = TRUE)/sqrt(length(imd_band_num[!is.na(imd_band_num)])),
            
            pass.mean = mean(pass_fail,na.rm = TRUE),
            sem.pass.mean = sd(pass_fail,na.rm = TRUE)/sqrt(length(pass_fail[!is.na(pass_fail)]))) %>%
  mutate(region = fct_reorder(region,pass.mean)) %>%
  left_join(sum.region.with[,c("region","with.mean","sem.with.mean")])  %>%
 left_join(imd.band) %>%
  mutate(Region = paste0(region," (IMD",imd_band,")")) %>%
  dplyr::rename(IMD_band = "imd_band")

p1=ggplot(sum.region.pass.fail,aes(x = pass.mean, y = with.mean,colour=IMD_band)) + geom_point()  + theme_classic() +
  geom_vline(xintercept = mean(combine.feat.unique$pass_fail,na.rm = TRUE),linetype="dotted",colour="grey") + 
   geom_hline(yintercept = mean(combine.feat.unique$withdraw,na.rm = TRUE),linetype="dotted",colour="grey") + 
  geom_text_repel(aes(label = Region),show.legend = FALSE,size=3)  + labs(x = "Success rate", y = "Dropout rate") + theme(legend.position = "bottom")
```

* I will use module BBB as reference for the module, it's around the mean of withdrawal

* The module with higheest prop of withdrawal are CCC/DDD and the ones with less withdrawal are AAA and GGG. The socio econoic area does not seem to be too heterogenous apart from module AAA.

```{r exploratory-module-region}
sum.module.with <- combine.feat.unique %>%
  group_by(code_module) %>%
  summarise(with.mean = mean(withdraw,na.rm = TRUE),
            sem.with.mean = sd(withdraw,na.rm = TRUE)/sqrt(length(withdraw[!is.na(withdraw)]))) %>%
  ungroup() %>%
  mutate(code_module = fct_reorder(code_module,with.mean))

# Scotland, wales and ireland are the ones with lower levels of widthdrawal
sum.region.pass.fail <- combine.feat.unique %>%
  group_by(code_module) %>%
  summarise(pass.mean = mean(pass_fail,na.rm = TRUE),
            sem.pass.mean = sd(pass_fail,na.rm = TRUE)/sqrt(length(pass_fail[!is.na(pass_fail)])),
            imd.mean = round(mean(imd_band_num,na.rm = TRUE))) %>%
  mutate(code_module = fct_reorder(code_module,pass.mean)) %>%
  left_join(sum.module.with) %>%
   left_join(imd.band) %>%
  mutate(Module = paste0(code_module," (IMD",imd_band,")")) %>%
  dplyr::rename(IMD_band = imd_band)



p3=ggplot(sum.region.pass.fail,aes(x = pass.mean, y = with.mean,colour=IMD_band)) + geom_point()  + theme_classic() +
  geom_vline(xintercept = mean(combine.feat.unique$pass_fail,na.rm = TRUE),linetype="dotted",colour="grey") + 
   geom_hline(yintercept = mean(combine.feat.unique$withdraw,na.rm = TRUE),linetype="dotted",colour="grey") + 
  geom_text_repel(aes(label = Module),show.legend = FALSE,size=3)  + labs(x = "Success rate", y = "Dropout rate") + theme(legend.position = "bottom") 


pc1 = plot_grid(p1  + ggtitle("Success and dropout rate by region"),
              p3+ ggtitle("Success and dropout rate by module"),nrow=1)

```

### Module across years

```{r exploratory-with-module-by-year}
sum.module.with <- combine.feat.unique %>%
  group_by(code_module,semester,year) %>%
  summarise(with.mean = mean(withdraw,na.rm = TRUE),
            sem.with.mean = sd(withdraw,na.rm = TRUE)/sqrt(length(withdraw[!is.na(withdraw)]))) %>%
  ungroup() %>%
  mutate(code_module = fct_reorder(code_module,with.mean))

sum.region.pass.fail <- combine.feat.unique %>%
 group_by(code_module,semester,year) %>%
  summarise(pass.mean = mean(pass_fail,na.rm = TRUE),
            sem.pass.mean = sd(pass_fail,na.rm = TRUE)/sqrt(length(pass_fail[!is.na(pass_fail)]))) %>%
  ungroup() %>%
  mutate(code_module = fct_reorder(code_module,pass.mean)) %>%
  left_join(sum.module.with)


ggplot(sum.region.pass.fail,aes(x = pass.mean, y = with.mean,colour=code_module,shape=year)) + geom_point(stat = "identity",size=2)  + theme_bw() +
  geom_text_repel(aes(label = code_module)) + facet_wrap(~semester) +
  geom_hline(yintercept = mean(combine.feat.unique$withdraw,na.rm = TRUE),linetype="dotted")+
   geom_vline(xintercept = mean(combine.feat.unique$pass_fail,na.rm = TRUE),linetype="dotted")

ggsave("output/dropout-success-module-semester-year.png",width = 10,height = 10)
```

### Education

```{r exploratory-with-education-by-year}
# Scotland, wales and ireland are the ones with lower levels of widthdrawal
sum.edu.pass.fail <- combine.feat.unique %>%
  group_by(highest_education) %>%
  summarise(with.mean = mean(withdraw,na.rm = TRUE),
            sem.with.mean = sd(withdraw,na.rm = TRUE)/sqrt(length(withdraw[!is.na(withdraw)])),
            
    pass.mean = mean(pass_fail,na.rm = TRUE),
            sem.pass.mean = sd(pass_fail,na.rm = TRUE)/sqrt(length(pass_fail[!is.na(pass_fail)])),
    
            imd.mean = round(mean(imd_band_num,na.rm = TRUE))) %>%
  mutate(highest_education = fct_reorder(highest_education,pass.mean)) %>%
   left_join(imd.band) %>%
  mutate(Highest_education = paste0(highest_education," (IMD",imd_band,")")) %>%
  dplyr::rename(IMD_band = imd_band)


p4 = ggplot(sum.edu.pass.fail,aes(x = pass.mean, y = with.mean,colour=IMD_band)) + geom_point()  + theme_classic() +
  geom_vline(xintercept = mean(combine.feat.unique$pass_fail,na.rm = TRUE),linetype="dotted",colour="grey") + 
   geom_hline(yintercept = mean(combine.feat.unique$withdraw,na.rm = TRUE),linetype="dotted",colour="grey") + 
  geom_text_repel(aes(label = Highest_education),show.legend = FALSE,size=3)  + labs(x = "Success rate", y = "Dropout rate") + theme(legend.position = "bottom") 
```

### Gender

```{r exploratory-with-education-by-year}
# Scotland, wales and ireland are the ones with lower levels of widthdrawal
sum.edu.pass.fail <- combine.feat.unique %>%
  group_by(gender) %>%
  summarise(with.mean = mean(withdraw,na.rm = TRUE),
            sem.with.mean = sd(withdraw,na.rm = TRUE)/sqrt(length(withdraw[!is.na(withdraw)])),
            
    pass.mean = mean(pass_fail,na.rm = TRUE),
            sem.pass.mean = sd(pass_fail,na.rm = TRUE)/sqrt(length(pass_fail[!is.na(pass_fail)])),
    
            imd.mean = round(mean(imd_band_num,na.rm = TRUE))) %>%
  mutate(highest_education = fct_reorder(highest_education,pass.mean)) %>%
   left_join(imd.band) %>%
  mutate(Highest_education = paste0(highest_education," (IMD",imd_band,")")) %>%
  dplyr::rename(IMD_band = imd_band)


p4 = ggplot(sum.edu.pass.fail,aes(x = pass.mean, y = with.mean,colour=IMD_band)) + geom_point()  + theme_classic() +
  geom_vline(xintercept = mean(combine.feat.unique$pass_fail,na.rm = TRUE),linetype="dotted",colour="grey") + 
   geom_hline(yintercept = mean(combine.feat.unique$withdraw,na.rm = TRUE),linetype="dotted",colour="grey") + 
  geom_text_repel(aes(label = Highest_education),show.legend = FALSE,size=3)  + labs(x = "Success rate", y = "Dropout rate") + theme(legend.position = "bottom") 
```



## Exploratory: dropout/success by use of VLE 

* What's the frequency of final_results among people that clicked and not clicked?

```{r frequency-table-VLE-final-result,fig.height=5,fig.width=5}
combine.feat.clicks <- combine.feat.unique %>%
  left_join(click.students.activity) %>%
  mutate(final_result = factor(final_result, levels = c("Distinction","Pass",
                                                        "Fail","Withdrawn"))) 

freq.results <- combine.feat.clicks %>%
  group_by(id_student,final_result) %>%
  summarise(n.clicks = round(mean(n.clicks.act,na.rm = TRUE))) %>%
  mutate(VLE = ifelse(is.na(n.clicks),"Never used VLE","Used VLE")) %>%
  group_by(VLE,final_result) %>%
  summarise(n = n())

freq = ggplot(freq.results,aes(x = final_result,y = n, fill = VLE)) + geom_bar(stat = "identity",position = "dodge") + theme_classic() +
  geom_text(aes(label = round(n)), hjust=0.5, vjust=-1, size = 3,position=position_dodge(width=0.9)) + labs(x = "Final result", y = "Number of students")
```


```{r}
pc2 = plot_grid(p4,freq + theme(legend.position = "bottom"),nrow=1)
plot_grid(pc1,pc2,nrow=2,align = 'h', rel_widths = c(1, 1))

dir.create("output",showWarnings = FALSE)
ggsave("output/success-dropout-imb-region-module-edu.png",width = 10,height = 10)
ggsave("output/success-dropout-imb-region-module-edu.pdf",width = 10,height = 10)
```


```{r chi-square-results-vle}
freq.results <- combine.feat.clicks %>%
  group_by(id_student,final_result) %>%
  summarise(n.clicks = mean(n.clicks.act,na.rm = TRUE)) %>%
  mutate(VLE = ifelse(is.na(n.clicks),"Never used VLE","Used VLE"))

# Chi square
chisq.test(table(freq.results$final_result,freq.results$VLE))
```


```{r combined-clicks-ass}
#apply(combine.feat.clicks,2,function(x) sum(is.na(x)))

n.final.result <- combine.feat.unique %>%
  group_by(final_result) %>%
  summarise(n = n())

# sum all clicks done for each result and activity type and divide by number of students in each result category
result.clicks <- combine.feat.clicks %>%
  group_by(final_result,activity_type) %>%
  left_join(n.final.result) %>%
  summarise(sum.clicks = round(sum(n.clicks.act,na.rm = TRUE))/unique(n)) %>%
  filter(!is.na(activity_type)) %>%
  ungroup() %>%
  mutate(activity_type =  fct_reorder(activity_type,sum.clicks)) %>%
  mutate(final_result = factor(final_result, levels = c("Distinction","Pass",
                                                        "Fail","Withdrawn"))) 

ggplot(result.clicks,aes(x=final_result,y=sum.clicks,fill=activity_type)) + geom_bar(stat="identity",position = "dodge",colour="white") + theme_classic() + theme(legend.position = "bottom") + labs(x = "Final result", y = "Sum of clicks per activity type")+ geom_text(aes(label = round(sum.clicks)), hjust=0.5, vjust=-1, size = 2,position=position_dodge(width=0.9))
```


* quiz and outcontent most fruitful

```{r combined-clicks-ass-logistic}
# correlation
combine.feat.unique.num <- combine.feat.clicks %>%
  spread(key = activity_type, value = n.clicks.act) %>%
  mutate_at(.vars = c("oucontent","quiz","homepage","subpage","ouwiki","resource"),.funs = function(x) log(x)) %>%
  dplyr::select(pass_fail, oucontent,quiz,homepage,subpage,disability, forumng,semester ,year,gender,ouwiki,resource,code_module) %>%
  mutate_all(.funs = function(x) as.numeric(as.factor(x)))

corrplot::corrplot(cor(combine.feat.unique.num,use="pairwise.complete.obs"))

combine.feat.unique.num <- combine.feat.clicks %>%
  spread(key = activity_type, value = n.clicks.act)

apply(combine.feat.unique.num,2,function(x) sum(is.na(x)))

# remove homepage
m <- glm(pass_fail ~  disability + log(oucontent) + semester + year + gender + code_module, data = combine.feat.unique.num, family = binomial)

car::vif(m)

tidym <- tidy(m,conf.int=TRUE) 

tidym <- tidym %>%
  mutate(term = factor(term,levels = term)) %>%
  filter(term != "(Intercept)")

ggplot(tidym,aes(x = term, y = estimate,colour=-log10(p.value))) + geom_point() + geom_errorbar(aes(ymin = conf.low, ymax = conf.high))  +
  theme_bw() + geom_hline(yintercept = 0, linetype = "dotted") + coord_flip() + ggtitle("Clicks associated with success vs failure")
```


```{r exploratory-clicks-activity-module}
sum.region <- combine.feat.clicks %>%
  group_by(code_module,semester) %>%
  summarise(clicks.mean = mean(n.clicks.act,na.rm = TRUE),
            sem.clicks.mean = sd(n.clicks.act,na.rm = TRUE)/sqrt(length(n.clicks.act[!is.na(n.clicks.act)])),
            
            with.mean = mean(withdraw),
            sem.with.mean = sd(withdraw)/sqrt(length(withdraw))) %>%
  ungroup() %>%
  mutate(code_module = fct_reorder(code_module,with.mean))

ggplot(sum.region,aes(x = with.mean, y = clicks.mean,colour=code_module)) + geom_point(stat = "identity")  + theme_bw() +
  geom_errorbar(aes(ymin = clicks.mean - sem.clicks.mean, 
                    ymax = clicks.mean + sem.clicks.mean)) +
  geom_errorbarh(aes( xmin = with.mean - sem.with.mean, 
                    xmax = with.mean + sem.with.mean)) +
  geom_label_repel(aes(label = code_module)) + facet_wrap(~semester) + labs(x="Dropout rate",y="Mean of clicks")
```


# Logistic on withdral including both B and J together

* Mess are less likely to withdraw (does this depend on the region?)
* East anglian region as reference
* I am going to randomly take one observation from students with multiple courses

## Combined semester

```{r very-skewed-vars}
# centered on the mean
combine.feat.unique$comb.score.std <- scale(combine.feat.unique$comb.score,center = TRUE,scale = TRUE) 
combine.feat.unique$total.clicks.std <- scale(log(combine.feat.unique$total.clicks),center = TRUE,scale = TRUE)
combine.feat.unique$studied_credits.std <- scale(log(combine.feat.unique$studied_credits),center = TRUE,scale = TRUE)

par(mfrow=c(2,2))
hist(combine.feat.unique$comb.score.std)
hist(combine.feat.unique$total.clicks.std)
hist(combine.feat.unique$studied_credits.std)

# Recode modules
regions <- names(table(combine.feat.unique$region))[!(names(table(combine.feat.unique$region)) %in% "South West Region")]
combine.feat.unique <- combine.feat.unique %>%
  mutate(code_module = factor(code_module, 
                              levels = c("BBB","AAA","CCC","DDD","EEE","FFF","GGG")),
        
          region = factor(region, levels=c("South West Region",regions)))
```


* Very bad collinearity

```{r logistic1-withdraw}
# No imd
m <- glm(withdraw ~ gender + region + highest_education + studied_credits.std + disability +
    module_presentation_length + code_module + semester + year + total.clicks.std + n.students + num_of_prev_attempts_dich + comb.score.std, data = combine.feat.unique, family = binomial)

car::vif(m)
```

A few independent variables show quite high VIF which is symptoms of  multicollinearity. This makes sense because the number of students and length of a course are correlated with the module. 

```{r nstudents-by-module}
ggplot(combine.feat.unique,aes(x= code_module, y = n.students,fill=semester)) + geom_bar(stat = "identity",position = "dodge") + theme_bw()
```

### Remove `n.students` and `module_presentation_length`

```{r logistic2-withdraw}
m <- glm(withdraw ~ gender + region + highest_education + studied_credits.std + disability + code_module + semester + year + total.clicks.std +  num_of_prev_attempts_dich + comb.score.std + imd_band_num, data = combine.feat.unique, family = binomial)

car::vif(m) # looks better

summary(m)

tidym.combined <- tidy(m,conf.int=TRUE) %>%
  mutate(term = factor(term,levels = term)) %>%
  filter(term != "(Intercept)")
```


```{r plot-logistic-with}
ggplot(tidym.combined,aes(x = term, y = estimate)) + geom_point() + geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = "dotted") + coord_flip() + ggtitle("Factors associated with withdrawal")
```


### Diagnostics and prediction

```{r check-predictions}
aug1 <- cbind(combine.feat.unique, predict(m, newdata = combine.feat.unique, type = "link",
    se = TRUE))
pairs(aug1[,c("fit","total.clicks.std","studied_credits.std")])


aug1 <- cbind(combine.feat.unique, predict(m, newdata = combine.feat.unique, type = "response",
    se = TRUE))
ggplot(aug1,aes(x = fit , fill = factor(withdraw))) + geom_histogram(alpha=0.5)+ theme_bw()

# Chisquare t-test
HLgof.test(fit = aug1$fit[!is.na(aug1$fit)], 
           obs = aug1$withdraw[!is.na(aug1$fit)])


# Roc curve
aug1 <- cbind(combine.feat.unique, predict(m, newdata = combine.feat.unique, type = "response",
    se = TRUE))
roc.log <- ggplot(aug1, aes(d = withdraw, m = fit)) + geom_roc() + theme_bw()
roc.log + 
  ggtitle("Withdrawal - combined semesters") + 
  annotate("text", x = .75, y = .25, 
           label = paste("AUC =", round(calc_auc(roc.log)$AUC, 2))) +
  geom_abline(sslope = 1, intercept = 0, linetype = "dotted") +
  scale_x_continuous("1 - Specificity", breaks = seq(0, 1, by = .1))
```


# Logistic on success including both B and J together

## Combined semester

```{r logistic-success}
m <- glm(pass_fail ~ gender + region + highest_education + studied_credits.std + disability + code_module + semester + year + total.clicks.std +  num_of_prev_attempts_dich + comb.score.std + imd_band_num, data = combine.feat.unique, family = binomial)

car::vif(m) # looks better

summary(m)

tidym.combined <- tidy(m,conf.int=TRUE) %>%
  mutate(term = factor(term,levels = term)) %>%
  filter(term != "(Intercept)")
```


```{r plot-logistic-success}
ggplot(tidym.combined,aes(x = term, y = estimate)) + geom_point() + geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = "dotted") + coord_flip() + ggtitle("Factors associated with withdrawal")
```


## Diagnostics and prediction

```{r check-predictions-success}
aug1 <- cbind(combine.feat.unique, predict(m, newdata = combine.feat.unique, type = "link",
    se = TRUE))
pairs(aug1[,c("fit","total.clicks.std","studied_credits.std")])


aug1 <- cbind(combine.feat.unique, predict(m, newdata = combine.feat.unique, type = "response",
    se = TRUE))
ggplot(aug1,aes(x = fit , fill = factor(withdraw))) + geom_histogram(alpha=0.5)+ theme_bw()

# Chisquare t-test
HLgof.test(fit = aug1$fit[!is.na(aug1$fit)], 
           obs = aug1$withdraw[!is.na(aug1$fit)])


# Roc curve
aug1 <- cbind(combine.feat.unique, predict(m, newdata = combine.feat.unique, type = "response",
    se = TRUE))
roc.log <- ggplot(aug1, aes(d = withdraw, m = fit)) + geom_roc() + theme_bw()
roc.log + 
  ggtitle("Withdrawal - combined semesters") + 
  annotate("text", x = .75, y = .25, 
           label = paste("AUC =", round(calc_auc(roc.log)$AUC, 2))) +
  geom_abline(sslope = 1, intercept = 0, linetype = "dotted") +
  scale_x_continuous("1 - Specificity", breaks = seq(0, 1, by = .1))
```



# Withdrawal by semester

## Semester B

* Remove regions, it was adding coplexity to the model but not really making that much of a difference. Should probably test it.

```{r}
semB <- subset(combine.feat.unique, semester %in% "B")
m <- glm(withdraw ~ gender + region + highest_education + studied_credits.std + disability + code_module + year + total.clicks.std +  num_of_prev_attempts_dich + comb.score.std + imd_band_num, data = semB, family = binomial)

car::vif(m) # oooks better

summary(m)

tidyB <- tidy(m,conf.int=TRUE) %>%
  mutate(term = factor(term,levels = term)) %>%
  filter(term != "(Intercept)")

ggplot(tidyB,aes(x = term, y = estimate)) + geom_point() + geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = "dotted") + coord_flip() + ggtitle("Factors associated with withdrawal")
```

```{r}
aug1 <- cbind(semB, predict(m, newdata = semB, type = "response",
    se = TRUE))

ggplot(aug1,aes(x = fit , fill = factor(withdraw))) + geom_histogram(alpha=0.5)+ theme_bw()

HLgof.test(fit = aug1$fit[!is.na(aug1$fit)], 
           obs = aug1$withdraw[!is.na(aug1$fit)])
```

## Semester J

```{r}
semJ <- subset(combine.feat.unique, semester %in% "J")
m <- glm(withdraw ~ gender + region + highest_education + studied_credits.std + disability + code_module + year + total.clicks.std +  num_of_prev_attempts_dich + comb.score.std + imd_band_num, data = semJ, family = binomial)

car::vif(m) # oooks better

summary(m)

tidyJ <- tidy(m,conf.int=TRUE) %>%
  mutate(term = factor(term,levels = term)) %>%
  filter(term != "(Intercept)")

ggplot(tidyJ,aes(x = term, y = estimate)) + geom_point() + geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = "dotted") + coord_flip() + ggtitle("Factors associated with withdrawal - J")
```

```{r}
aug1 <- cbind(semJ, predict(m, newdata = semJ, type = "response",
    se = TRUE))

ggplot(aug1,aes(x = fit , fill = factor(withdraw))) + geom_histogram(alpha=0.5)+ theme_bw()

HLgof.test(fit = aug1$fit[!is.na(aug1$fit)], 
           obs = aug1$withdraw[!is.na(aug1$fit)])
```


## Compare B and J: difference of estimates

```{r blue-diff-betas}
tidyB <- tidyB %>%
  mutate(semester = "B")

tidyJ <- tidyJ %>%
  mutate(semester = "J")

# Differences of betas
tidy.compare.beta <- bind_rows(tidyB,tidyJ) %>%
  dplyr::select(term,estimate,semester) %>%
  spread(key = semester, value = estimate)  %>%
  mutate(diff.estimate = B - J)

# Uncertainty of difference
tidy.compare.std <- bind_rows(tidyB,tidyJ) %>%
  dplyr::select(term,std.error,semester) %>%
  spread(key = semester, value = std.error )  %>%
  group_by(term) %>%
  mutate(diff.std = sqrt(B^2 + J^2)) %>%
  dplyr::select(-B,-J)

# Combine and CI
tidy.compare <- tidy.compare.beta %>%
  left_join(tidy.compare.std) %>%
  mutate(diff.estimate =  diff.estimate/diff.std) %>%
  mutate(conf.low = diff.estimate - 1.96*diff.std,
         conf.high = diff.estimate + 1.96*diff.std) 

ggplot(tidy.compare,aes(x=term,y=diff.estimate)) + geom_point() +
  geom_errorbar(aes(ymin = conf.low,ymax= conf.high),width=0.5) + theme_bw() + geom_hline(yintercept = 0,linetype="dotted") + coord_flip() + labs(y = "B - J") + ggtitle("Differences between semesters")
```


# Linear model on score - not used

Now I will study what factors influence the total score obtained with assignments and exams.

It is slightly skewed. 

```{r}
hist(combine.feat.unique$comb.score)
```

## Combined by semester

```{r}
m <- lm(comb.score ~ gender + region + highest_education + studied_credits.std + disability + code_module + semester + year + total.clicks.std +  num_of_prev_attempts_dich + final_result, data = combine.feat.unique)

car::vif(m)

par(mfrow=c(2,2))
plot(m)

```

* diagnostics

```{r}
library(MASS)

qqnorm(combine.feat.unique$comb.score,
       ylab="Quantiles for comb.score")
qqline(combine.feat.unique$comb.score, 
       col="red")

Box = boxcox(combine.feat.unique$comb.score + 1 ~ 1,              # Transform Turbidity as a single vector
             lambda = seq(-6,6,0.1))

Cox = data.frame(Box$x, Box$y)            # Create a data frame with the results

Cox2 = Cox[with(Cox, order(-Cox$Box.y)),] # Order the new data frame by decreasing y

Cox2[1,]                                  # Display the lambda with the greatest
                                          #    log likelihood

lambda = Cox2[1, "Box.x"]                 # Extract that lambda

combine.feat.unique$comb.score.box = (combine.feat.unique$comb.score ^ lambda - 1)/lambda   # Transform the original data

combine.feat.unique$comb.score.box.st <- scale(combine.feat.unique$comb.score.box,center = TRUE,scale = TRUE)

qqnorm(combine.feat.unique$comb.score.box,
       ylab="Quantiles for comb.score")
qqline(combine.feat.unique$comb.score.box, 
       col="red")


```



```{r}
m <- lm(comb.score.std ~ gender + region + highest_education + studied_credits.std + disability + code_module + semester + year + total.clicks.std +  num_of_prev_attempts_dich, data = combine.feat.unique)

par(mfrow=c(2,4))
plot(m)

# After box-cox
m <- lm(comb.score.box.st ~ gender + region + highest_education + studied_credits.std + disability + code_module + semester + year + total.clicks.std +  num_of_prev_attempts_dich, data = combine.feat.unique)

plot(combine.feat.unique$comb.score.box.st,combine.feat.unique$comb.score)

plot(m)

car::vif(m)


tidym <- tidy(m,conf.int=TRUE) 

tidym <- tidym %>%
  mutate(term = factor(term,levels = term)) %>%
  filter(term != "(Intercept)")

ggplot(tidym,aes(x = term, y = estimate,colour=-log10(p.value))) + geom_point() + geom_errorbar(aes(ymin = conf.low, ymax = conf.high))  +
  theme_bw() + geom_hline(yintercept = 0, linetype = "dotted") + coord_flip() + ggtitle("Factors associated with total score")
```


## Semester B

```{r semesterB-total-score}
m <- lm(comb.score.std ~ gender + region + highest_education + studied_credits.std + disability + code_module + year + total.clicks.std +  num_of_prev_attempts_dich, data = combine.feat.unique[combine.feat.unique$semester %in% "B",])

par(mfrow=c(2,4))
plot(m)

# After box-cox
m <- lm(comb.score.box.st ~ gender + region + highest_education + studied_credits.std + disability + code_module  + year + total.clicks.std +  num_of_prev_attempts_dich, data = combine.feat.unique[combine.feat.unique$semester %in% "B",])

par(mfrow=c(2,2))
plot(m)

car::vif(m)
```

* From the plot of residuals and quite high correlation between the y variable and the residuals it looks like we are not explaining that much and indeed the Rsquared of the model is just 0.2.  

```{r residuals-B-sem}
aug <- augment(m)

pairs(aug[,c(".resid",".fitted","total.clicks.std",".std.resid","studied_credits.std","comb.score.box.st")])

ggplot(aug,aes(x = code_module,y=.resid,fill=region)) + geom_boxplot()
```


```{r}
# After box-cox
m <- lm(comb.score.box.st ~ gender + highest_education + total.clicks.std +  num_of_prev_attempts_dich, data = combine.feat.unique[combine.feat.unique$semester %in% "B",])

summary(m)

par(mfrow=c(2,2))
plot(m)

aug <- augment(m)

pairs(aug[,c(".resid",".fitted","total.clicks.std",".std.resid","comb.score.box.st")])
```

## Semester J

```{r semesterJ-total-score}
m <- lm(comb.score.std ~ gender + region + highest_education + studied_credits.std + disability + code_module + year + total.clicks.std +  num_of_prev_attempts_dich, data = combine.feat.unique[combine.feat.unique$semester %in% "J",])

summary(m)

par(mfrow=c(2,2))
plot(m)

# After box-cox
m <- lm(comb.score.box.st ~ gender + highest_education + studied_credits.std + disability + code_module  + year + total.clicks.std +  num_of_prev_attempts_dich, data = combine.feat.unique[combine.feat.unique$semester %in% "J",])

par(mfrow=c(2,2))
plot(m)

summary(m)

car::vif(m)
```

* Same thing for this semester 

```{r residuals-J-sem}
aug <- augment(m)

pairs(aug[,c(".resid",".fitted","total.clicks.std",".std.resid","studied_credits.std","comb.score.box.st")])
```


```{r}
purl("nous-student.Rmd", output = "nous-student.R", documentation = 2)
```


